---
title: "Bay Study Fish Data Summary"
author: "Dave Bosworth"
date: "3/10/2020"
output: 
  html_document: 
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose
The heat map plots in this document provide summaries of the number of surveys conducted at each station during each season for the Bay Study midwater trawl. Each Series, which is a group of stations, are plotted separately. Series 1 stations are considered the "core" sampling locations, and Series 2-4 stations were established later in the study period.

# Import and Prepare Data

Load R packages:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
```

Import data:
```{r import data}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/DWR - Seasonal monitoring report - Documents/Data"
  )
)  

# Import fish data
load(file = paste0(sharepoint_path, "/BayStudyFish.RData"))
```

Prepare midwater trawl data for plots:
```{r prepare mwt data, message = FALSE}
mwt_clean <- midwater_trawl_data %>% 
  # adjust years for surveys 1 and 2 (January and March)
  mutate(year_index = if_else(Survey <= 2, Year - 1, Year)) %>% 
  # don't include year_index 1979
  filter(year_index != 1979) %>% 
  # add a season variable
  mutate(
    season = case_when(
      Survey == 12 ~ "Winter",
      Survey >= 9 ~ "Fall",
      Survey >= 6 ~ "Summer",
      Survey >= 3 ~ "Spring",
      TRUE ~ "Winter"
    )
  ) %>% 
  # summarize counts by Series, Station, year_index, and season
  count(Series, Station, year_index, season) %>% 
  rename(num_survey = n) %>% 
  # convert Station and season to factors
  mutate(
    Station = factor(Station),
    season = factor(season, levels = c("Winter", "Spring", "Summer", "Fall"))
  )

# Create a nested df with year_index = 2018 to fill in zeros
mwt_2018 <- mwt_clean %>% 
  group_nest(Series) %>% 
  mutate(
    data1 = map(
      data, 
      ~slice(.x, 1) %>% 
      mutate(year_index = 2018, num_survey = 0)
    )
  ) %>% 
  select(-data)

# Fill in zeros for missing Station, Year, Season combos
mwt_complete <- mwt_clean %>% 
  group_nest(Series) %>% 
  left_join(mwt_2018) %>%
  mutate(
    data_all = map2(
      data, 
      data1, 
      ~bind_rows(.x, .y) %>% 
      mutate(Station = fct_drop(Station)) %>% 
      complete(Station, year_index, season, fill = list(num_survey = 0)) 
    )
  ) %>% 
  select(-c(data, data1))
```

# Midwater Trawl Summary

Create function for heat map plots
```{r fun heatmap}
heatmap_summ <- function(df) {
  p <- df %>% 
    ggplot(
      aes(
        x = year_index,
        y = Station
      )
    ) +
    geom_tile(aes(fill = num_survey)) +
    facet_grid(cols = vars(season)) +
    scale_fill_viridis_c(
      name = "# of surveys",
      direction = -1
    ) +
    xlab("Year") +
    scale_x_continuous(
      limits = c(1979.4, 2018.6),
      breaks = c(seq(1980, 2015, 5), 2018)
    ) +
    theme_light() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1),
      strip.text = element_text(color = "black")
    )
  
  return(p)
}
```

Run heatmap_summ function on each Series
```{r create heatmap plots}
mwt_plots <- mwt_complete %>% 
  mutate(heatmap_plot = map(data_all, .f = heatmap_summ))
```

## Series 1 stations
```{r plot series 1, fig.height = 7, fig.width = 9}
mwt_plots %>% filter(Series == 1) %>% pull(heatmap_plot) %>% chuck(1)
```

## Series 2 stations
```{r plot series 2, fig.height = 4, fig.width = 9}
mwt_plots %>% filter(Series == 2) %>% pull(heatmap_plot) %>% chuck(1)
```

## Series 3 stations
```{r plot series 3, fig.height = 3, fig.width = 9}
mwt_plots %>% filter(Series == 3) %>% pull(heatmap_plot) %>% chuck(1)
```

## Series 4 stations
```{r plot series 4, fig.height = 3, fig.width = 9}
mwt_plots %>% filter(Series == 4) %>% pull(heatmap_plot) %>% chuck(1)
```
