---
title: "Northern Anchovy Plot Comparison"
author: "Dave Bosworth"
date: "3/10/2020"
output: 
  html_document: 
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose
This document compares plots of the Bay Study northern anchovy CPUE including all stations versus only the Series 1 stations. Series 1 stations are considered the "core" sampling locations.

# Import and Prepare Data

Load R packages:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(scales)
library(smonitr)
```

Import data:
```{r import data}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/DWR - Seasonal monitoring report - Documents/Data"
  )
)  

# Import fish data
load(file = paste0(sharepoint_path, "/BayStudyFish.RData"))
```

Prepare northern anchovy data for plots:
```{r prepare noranc data}
# Clean and modify midwater_trawl_data df
noranc <- midwater_trawl_data %>% 
  select(Year:ChanShoal, TowVolume, NORANC) %>% 
  # Used Survey = 6, 7, 8 instead of month to include complete dataset
  filter(Survey >= 6 & Survey <= 8) %>% 
  # remove data from July 2016 due to invalid tow data
  filter(!(Year == 2016 & Survey == 7))

# Calculate CPUE values & average them for each year
  # All stations
  noranc_cpue_all <- noranc %>% 
    mutate(cpue = (NORANC/TowVolume) * 10000) %>% 
    group_by(Year) %>% 
    summarize(ave_cpue = mean(cpue)) %>% 
    ungroup() %>% 
    mutate(Year = factor(Year))
  
  # Just Series 1 stations
  noranc_cpue_s1 <- noranc %>% 
    filter(Series == 1) %>% 
    mutate(cpue = (NORANC/TowVolume) * 10000) %>% 
    group_by(Year) %>% 
    summarize(ave_cpue = mean(cpue)) %>% 
    ungroup() %>% 
    mutate(Year = factor(Year))
```

# Create Plot functions

Define global settings for plots:
```{r global plot settings}
# Define report year
report_year <- 2018

# Calculate long-term average of CPUE
# All Stations
lt_avg_cpue_all <- mean(noranc_cpue_all$ave_cpue)

# Just Series 1 stations
lt_avg_cpue_s1 <- mean(noranc_cpue_s1$ave_cpue)
```

Base plot function:
```{r fun base plot}
base_plot <- function(df, ltavg) {
  p <- df %>% 
    ggplot(
      aes(
        x = Year,
        y = ave_cpue
      )
    ) +
    geom_col() +
    # apply custom theme
    theme_smr() +
    # customize axis labels
    ylab(expression(paste("Average CPUE (fish/10,000m"^{3}, ")"))) +
    std_x_axis_label("summer") +
    # add horizontal line for long-term average CPUE
    lt_avg_line(ltavg)
  
  return(p)
}
```

Function to add options for plots of all years:
```{r fun all yrs plot}
plot_all_yrs <- function() {
  list(
    # define y-axis limits, breaks and add thousanths comma
    scale_y_continuous(
      limits = c(0, 1540),
      breaks = seq(0, 1500, by = 250),
      labels = label_comma()
    ),
    # standardize x-axis
    std_x_axis_all_years(report_year, "discrete", start_yr = 1980),
    # add markers for missing data
    missing_data_symb(noranc_cpue_all, Year, report_year, 2)
  )
}
```

Function to add options for plots of recent years:
```{r fun rec yrs plot}
plot_rec_yrs <- function() {
  list(
    # define y-axis limits, breaks and add thousanths comma
    scale_y_continuous(
      limits = c(0, 740),
      breaks = seq(0, 700, by = 100),
      labels = label_comma()
    ),
    # standardize x-axis
    std_x_axis_rec_years(report_year, "discrete"),
    # add markers for missing data
    missing_data_symb(noranc_cpue_all, Year, report_year, 4)
  )
}
```

# Plots of all years

## All Stations

```{r plot all yrs all stations, fig.height = 4, fig.width = 6}
base_plot(noranc_cpue_all, lt_avg_cpue_all) + plot_all_yrs()
```

## Just Series 1 Stations

```{r plot all yrs s1 stations, fig.height = 4, fig.width = 6}
base_plot(noranc_cpue_s1, lt_avg_cpue_s1) + plot_all_yrs()
```

# Plots of recent years

## All Stations

```{r plot rec yrs all stations, warning = FALSE, fig.height = 4, fig.width = 6}
base_plot(noranc_cpue_all, lt_avg_cpue_all) + plot_rec_yrs()
```

## Just Series 1 Stations

```{r plot rec yrs s1 stations, warning = FALSE, fig.height = 4, fig.width = 6}
base_plot(noranc_cpue_s1, lt_avg_cpue_s1) + plot_rec_yrs()
```


